---
title: "Group 1 – Survival Analysis Assignment 2"
author: "Group 1"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_depth: 3
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE,
                      fig.width = 8, fig.height = 5)
```



```{r packages-data}
library(survival)
library(dplyr)
library(tidyr)
library(finalfit)
library(knitr)
library(kableExtra)
library(ggplot2)
library(survminer)
library(flexsurv)
library(gridExtra)
library(tibble)
library(ggpubr)
library(muhaz)
library(splines)
library(rms)
library(patchwork)
library(purrr)
library(forcats)
```




```{r}

data(cancer, package = "survival")
df_veteran <- veteran

if (all(sort(unique(na.omit(df_veteran$status))) == c(1, 2))) {
  df_veteran <- df_veteran %>% mutate(status = status - 1)
}

df_veteran <- df_veteran %>%
  mutate(
    trt = factor(trt, levels = c(1, 2), labels = c("Standard", "Test")),
    celltype = as.character(celltype),
    celltype = case_when(
      celltype %in% c("1", "squamous") ~ "Squamous",
      celltype %in% c("2", "smallcell") ~ "Small cell",
      celltype %in% c("3", "adeno")     ~ "Adeno",
      celltype %in% c("4", "large")     ~ "Large",
      TRUE ~ celltype
    ),
    celltype = factor(celltype, levels = c("Squamous", "Small cell", "Adeno", "Large")),
    prior = factor(ifelse(prior == 0, "No", "Yes"), levels = c("No", "Yes"))
  )

```



### Survival object

```{r}
surv_obj <- Surv(time = df_veteran$time, event = df_veteran$status)
```


# Step 1: Kaplan-Meier Estimator



```{r km-overall-fit}

#********* Kaplan-Meier fit
km_overall <- survfit(surv_obj ~ 1, data = df_veteran)


#********* Stratified empirical KM estimate by treatment
km_trt <- survfit(surv_obj ~ trt, data = df_veteran)
km_trt_sum<-summary(km_trt)

```



## Kaplan-Meier Emprical Estimates

```{r}

km_table <- data.frame(
  Group = gsub("trt=", "", km_trt_sum$strata),
  time = km_trt_sum$time,
  n.risk = km_trt_sum$n.risk,
  n.event = km_trt_sum$n.event,
  survival = km_trt_sum$surv,
  lower = km_trt_sum$lower,
  upper = km_trt_sum$upper
) %>%
  mutate(
    survival = round(survival, 3),
    lower    = round(lower, 3),
    upper    = round(upper, 3),
    CI = paste0(lower, " - ", upper)
  ) %>%
  select(Group, time, n.risk, n.event, survival,  CI)

km_standard <- km_table %>%
  filter(Group == "Standard") %>%
  select(-Group) %>%
  slice(1:15)

standard_tbl <- kbl(
  km_standard,
  format = "html",
  align = "lccccc",
  caption = ""
) %>%
  kable_styling(
    full_width = FALSE,
    font_size = 18,
    bootstrap_options = c("striped", "hover", "condensed", "responsive")
  ) %>%
  row_spec(0, bold = TRUE, color = "white", background = "#1F4E79") %>%
  row_spec(1:nrow(km_standard), extra_css = "border-bottom: 1px solid #999;") %>%
  column_spec(1, color = "black", width = "8em") %>%
  column_spec(2, color = "black", width = "8em") %>%
  column_spec(3, color = "black", width = "8em") %>%
  column_spec(4, color = "black", width = "8em") %>%
  column_spec(5, color = "black", width = "8em") 

standard_tbl

save_kable(standard_tbl, "km_standard_table.png", zoom = 4)
```








$$\textbf{Test group}$$



```{r}
km_test <- km_table %>%
  filter(Group == "Test") %>%
  select(-Group) %>%
  slice(1:15)

test_tbl <- kbl(
  km_test,
  format = "html",
  align = "lccccc",
  caption = ""
) %>%
  kable_styling(
    full_width = FALSE,
    font_size = 18,
    bootstrap_options = c("striped", "hover", "condensed", "responsive")
  ) %>%
  row_spec(0, bold = TRUE, color = "white", background = "#1F4E79") %>%
  row_spec(1:nrow(km_test), extra_css = "border-bottom: 1px solid #999;") %>%
  column_spec(1, color = "black", width = "8em") %>%
  column_spec(2, color = "black", width = "8em") %>%
  column_spec(3, color = "black", width = "8em") %>%
  column_spec(4, color = "black", width = "8em") %>%
  column_spec(5, color = "black", width = "8em")

test_tbl

save_kable(test_tbl, "km_test_table.png", zoom = 4)
```








## 1.1 & 1.2 Percentile



```{r}

q_overall <- quantile(km_overall, probs = c(0.25, 0.50, 0.75))
q_trt     <- quantile(km_trt, probs = c(0.25, 0.50, 0.75))

#******** Overall table
overall_tbl <- data.frame(
  Group = "Overall",
  check.names = FALSE,
  `25%` = paste0(q_overall$quantile["25"], " (", q_overall$lower["25"], " - ", q_overall$upper["25"], ")"),
  `50%` = paste0(q_overall$quantile["50"], " (", q_overall$lower["50"], " - ", q_overall$upper["50"], ")"),
  `75%` = paste0(q_overall$quantile["75"], " (", q_overall$lower["75"], " - ", q_overall$upper["75"], ")")
)

#******* Treatment table
trt_tbl <- data.frame(
  Group = c("Standard", "Test"),
  check.names = FALSE,
  `25%` = paste0(q_trt$quantile[, "25"], " (", q_trt$lower[, "25"], " - ", q_trt$upper[, "25"], ")"),
  `50%` = paste0(q_trt$quantile[, "50"], " (", q_trt$lower[, "50"], " - ", q_trt$upper[, "50"], ")"),
  `75%` = paste0(q_trt$quantile[, "75"], " (", q_trt$lower[, "75"], " - ", q_trt$upper[, "75"], ")")
)


final_tbl <- bind_rows(overall_tbl, trt_tbl)


final_kbl <- kbl(
  final_tbl,
  format = "html",
  align = "lccc",
  caption = ""
) %>%
  kable_styling(
    full_width = FALSE,
    font_size = 18,
    bootstrap_options = c("striped", "hover", "condensed", "responsive")
  ) %>%
  row_spec(0, bold = TRUE, color = "white", background = "#1F4E79") %>%
  row_spec(1:nrow(final_tbl), extra_css = "border-bottom: 1px solid #999;") %>%
  column_spec(1, color = "black", width = "10em") %>%
  column_spec(2:4, color = "black", width = "12em")


final_kbl

save_kable(final_kbl, "percentile.png", zoom = 4)
```



## 1.3 Kaplan-Meier curve by treatment group



```{r}

df_veteran$trt <- factor(df_veteran$trt, levels = c("Standard", "Test"))


surv_obj <- Surv(df_veteran$time, df_veteran$status)

#******** Kaplan-Meier fit
km_trt <- survfit(surv_obj ~ trt, data = df_veteran)

#******** KM percentiles
q_trt <- quantile(km_trt, probs = c(0.25, 0.50, 0.75))

std_25  <- q_trt$quantile["trt=Standard", "25"]
std_50  <- q_trt$quantile["trt=Standard", "50"]
std_75  <- q_trt$quantile["trt=Standard", "75"]

test_25 <- q_trt$quantile["trt=Test", "25"]
test_50 <- q_trt$quantile["trt=Test", "50"]
test_75 <- q_trt$quantile["trt=Test", "75"]


#********** Generalized Gamma model

gg_model <- flexsurvreg(
  Surv(time, status) ~ trt,
  data = df_veteran,
  dist = "gengamma"
)

tt <- seq(0, max(df_veteran$time, na.rm = TRUE), length.out = 400)

pred_std <- summary(
  gg_model,
  newdata = data.frame(trt = factor("Standard", levels = levels(df_veteran$trt))),
  t = tt,
  type = "survival",
  ci = FALSE
)[[1]]

pred_test <- summary(
  gg_model,
  newdata = data.frame(trt = factor("Test", levels = levels(df_veteran$trt))),
  t = tt,
  type = "survival",
  ci = FALSE
)[[1]]

gen_std_df <- data.frame(
  time = pred_std$time,
  surv = pred_std$est,
  curve = "GenGamma Standard"
)

gen_test_df <- data.frame(
  time = pred_test$time,
  surv = pred_test$est,
  curve = "GenGamma Test"
)

gen_df <- rbind(gen_std_df, gen_test_df)


#******** KM plot

p <- ggsurvplot(
  km_trt,
  data = df_veteran,
  conf.int = TRUE,
  pval = TRUE,
  risk.table = TRUE,
  surv.median.line = "hv",
  xlab = "Time (days)",
  ylab = "Survival probability",
  legend.title = "Curve",
  legend.labs = c("KM Standard", "KM Test"),
  title = "Kaplan-Meier and Generalized Gamma Curves by Treatment",
  palette = c("#1F77B4", "darkgreen"),
  risk.table.col = "strata",
  ggtheme = theme_bw()
)


p$plot <- p$plot +
  geom_hline(yintercept = 0.75, linetype = "dashed", color = "darkgreen") +
  geom_hline(yintercept = 0.50, linetype = "dashed", color = "purple") +
  geom_hline(yintercept = 0.25, linetype = "dashed", color = "orange") +

  geom_segment(aes(x = std_25, xend = std_25, y = 0, yend = 0.75),
               linetype = "dotted", color = "#1F77B4") +
  geom_segment(aes(x = std_50, xend = std_50, y = 0, yend = 0.50),
               linetype = "dotted", color = "#1F77B4") +
  geom_segment(aes(x = std_75, xend = std_75, y = 0, yend = 0.25),
               linetype = "dotted", color = "#1F77B4") +

  geom_segment(aes(x = test_25, xend = test_25, y = 0, yend = 0.75),
               linetype = "dotted", color = "darkgreen") +
  geom_segment(aes(x = test_50, xend = test_50, y = 0, yend = 0.50),
               linetype = "dotted", color = "darkgreen") +
  geom_segment(aes(x = test_75, xend = test_75, y = 0, yend = 0.25),
               linetype = "dotted", color = "darkgreen") +

  geom_line(
    data = gen_df,
    aes(x = time, y = surv, color = curve),
    linewidth = 1.1,
    inherit.aes = FALSE
  ) +

  scale_color_manual(
    name = "Curve",
    values = c(
      "trt=Standard" = "#1F77B4",
      "trt=Test" = "darkgreen",
      "GenGamma Standard" = "black",
      "GenGamma Test" = "brown"
    ),
    breaks = c("trt=Standard", "trt=Test", "GenGamma Standard", "GenGamma Test"),
    labels = c("KM Standard", "KM Test", "GenGamma Standard", "GenGamma Test")
  ) +

  theme(
    plot.title = element_text(face = "bold", size = 16),
    axis.title = element_text(size = 14),
    legend.title = element_text(size = 12),
    legend.text = element_text(size = 11)
  )

p
```




## Survival curves grouped by the cell type


```{r}

df_veteran$celltype <- factor(df_veteran$celltype)
df_veteran$trt <- factor(df_veteran$trt, levels = c("Standard", "Test"))


plot_list <- lapply(levels(df_veteran$celltype), function(ct) {
  
  df_sub <- subset(df_veteran, celltype == ct)
  
  #***********  KM fit
  fit_sub <- survfit(Surv(time, status) ~ trt, data = df_sub)
  
  #********* Generalized gamma model within this cell type
  gg_model <- flexsurvreg(
    Surv(time, status) ~ trt,
    data = df_sub,
    dist = "gengamma"
  )
  
  tt <- seq(0, max(df_sub$time, na.rm = TRUE), length.out = 300)
  
  pred_std <- summary(
    gg_model,
    newdata = data.frame(trt = factor("Standard", levels = levels(df_veteran$trt))),
    t = tt,
    type = "survival",
    ci = FALSE
  )[[1]]
  
  pred_test <- summary(
    gg_model,
    newdata = data.frame(trt = factor("Test", levels = levels(df_veteran$trt))),
    t = tt,
    type = "survival",
    ci = FALSE
  )[[1]]
  
  gen_std_df <- data.frame(
    time = pred_std$time,
    surv = pred_std$est,
    curve = "GenGamma Standard"
  )
  
  gen_test_df <- data.frame(
    time = pred_test$time,
    surv = pred_test$est,
    curve = "GenGamma Test"
  )
  
  gen_df <- rbind(gen_std_df, gen_test_df)
  

  p <- ggsurvplot(
    fit_sub,
    data = df_sub,
    conf.int = TRUE,
    pval = TRUE,
    risk.table = FALSE,
    surv.median.line = "hv",
    xlab = "Time (days)",
    ylab = "Survival probability",
    title = ct,
    legend.title = "Curve",
    legend.labs = c("KM Standard", "KM Test"),
    palette = c("#1F77B4", "black"),
    ggtheme = theme_pubr()
  )
  
  p$plot <- p$plot +
    geom_line(
      data = gen_df,
      aes(x = time, y = surv, color = curve),
      linewidth = 1.1,
      inherit.aes = FALSE
    ) +
    scale_color_manual(
      name = "Curve",
      values = c(
        "KM Standard" = "#1F77B4",
        "KM Test" = "black",
        "GenGamma Standard" = "brown",
        "GenGamma Test" = "darkblue"
      ),
      breaks = c("KM Standard", "KM Test", "GenGamma Standard", "GenGamma Test")
    ) +
    theme(
      plot.title = element_text(size = 10, face = "bold"),
      axis.title = element_text(size = 9),
      axis.text = element_text(size = 8),
      legend.title = element_text(size = 9),
      legend.text = element_text(size = 8),
      legend.position = "bottom"
    )
  
  p
})


plot_only <- lapply(plot_list, function(x) x$plot)

ggarrange(
  plotlist = plot_only,
  ncol = 2,
  nrow = 2,
  common.legend = TRUE,
  legend = "bottom"
)
```



# Step 2. Nelson-Aalen Analysis

```{r}
na_trt <- survfit(surv_obj ~ trt, data = df_veteran)
na_trt_sum <- summary(na_trt)
```


### Nelson-Aalen cumulative hazard and estimate of the hazard function of standard group







```{r}
na_table <- data.frame(
  Group = gsub("trt=", "", na_trt_sum$strata),
  time = na_trt_sum$time,
  n.risk = na_trt_sum$n.risk,
  n.event = na_trt_sum$n.event,
  cumhaz = na_trt_sum$cumhaz,
  lower = na_trt_sum$std.chaz,
  upper = na_trt_sum$std.chaz
)

if (!is.null(na_trt_sum$lower) & !is.null(na_trt_sum$upper)) {
  na_table$lower <- na_trt_sum$lower
  na_table$upper <- na_trt_sum$upper
} else {
  na_table$lower <- NA
  na_table$upper <- NA
}

na_table <- na_table %>%
  mutate(
    hazard = n.event / n.risk,
    cumhaz = round(cumhaz, 3),
    hazard = round(hazard, 3),
    lower  = round(lower, 3),
    upper  = round(upper, 3),
    CI = ifelse(is.na(lower), "-", paste0(lower, " - ", upper))
  ) %>%
  select(Group, time, n.risk, n.event, cumhaz, hazard, CI)

na_standard <- na_table %>%
  filter(Group == "Standard") %>%
  select(-Group) %>%
  slice(1:15)

na_standard_tbl <- kbl(
  na_standard,
  format = "html",
  align = "lcccccc",
  caption = ""
) %>%
  kable_styling(
    full_width = FALSE,
    font_size = 18,
    bootstrap_options = c("striped", "hover", "condensed", "responsive")
  ) %>%
  row_spec(0, bold = TRUE, color = "white", background = "#1F4E79") %>%
  row_spec(1:nrow(na_standard), extra_css = "border-bottom: 1px solid #999;") %>%
  column_spec(1:6, color = "black", width = "8em")

na_standard_tbl
save_kable(na_standard_tbl, "na_standard_table.png", zoom = 4)
```
### Nelson-Aalen cumulative hazard and estimate of the hazard function of test therapy group



### Test group

```{r}
na_test <- na_table %>%
  filter(Group == "Test") %>%
  select(-Group) %>%
  slice(1:15)

na_test_tbl <- kbl(
  na_test,
  format = "html",
  align = "lccccc",
  caption = ""
) %>%
  kable_styling(
    full_width = FALSE,
    font_size = 18,
    bootstrap_options = c("striped", "hover", "condensed", "responsive")
  ) %>%
  row_spec(0, bold = TRUE, color = "white", background = "#1F4E79") %>%
  row_spec(1:nrow(na_test), extra_css = "border-bottom: 1px solid #999;") %>%
  column_spec(1:5, color = "black", width = "8em")

na_test_tbl
save_kable(na_test_tbl, "na_test_table.png", zoom = 4)
```
### Plots



```{r}


df_veteran$trt <- factor(df_veteran$trt, levels = c("Standard", "Test"))

df_veteran <- df_veteran %>%
  mutate(event = ifelse(status == 2, 1, status))

#******** Nelson-Aalen cumulative hazard

na_fit <- survfit(
  Surv(time, event) ~ trt,
  data = df_veteran,
  type = "fleming-harrington"
)

na_sum <- summary(na_fit)

na_table <- data.frame(
  Group   = gsub("trt=", "", na_sum$strata),
  time    = na_sum$time,
  n.risk  = na_sum$n.risk,
  n.event = na_sum$n.event,
  cumhaz  = na_sum$cumhaz
)

#*********** Generalized gamma model

gg_model <- flexsurvreg(
  Surv(time, event) ~ trt,
  data = df_veteran,
  dist = "gengamma"
)

tt <- seq(1, max(df_veteran$time, na.rm = TRUE), length.out = 400)

#******** Predictions for Standard
gg_std_surv <- summary(
  gg_model,
  newdata = data.frame(trt = factor("Standard", levels = levels(df_veteran$trt))),
  t = tt,
  type = "survival",
  ci = FALSE
)[[1]]

gg_std_haz <- summary(
  gg_model,
  newdata = data.frame(trt = factor("Standard", levels = levels(df_veteran$trt))),
  t = tt,
  type = "hazard",
  ci = FALSE
)[[1]]

#************** Predictions for Test
gg_test_surv <- summary(
  gg_model,
  newdata = data.frame(trt = factor("Test", levels = levels(df_veteran$trt))),
  t = tt,
  type = "survival",
  ci = FALSE
)[[1]]

gg_test_haz <- summary(
  gg_model,
  newdata = data.frame(trt = factor("Test", levels = levels(df_veteran$trt))),
  t = tt,
  type = "hazard",
  ci = FALSE
)[[1]]

gen_df <- bind_rows(
  data.frame(
    Group = "Standard",
    time = gg_std_surv$time,
    cumhaz = -log(pmax(gg_std_surv$est, 1e-8)),
    hazard = gg_std_haz$est
  ),
  data.frame(
    Group = "Test",
    time = gg_test_surv$time,
    cumhaz = -log(pmax(gg_test_surv$est, 1e-8)),
    hazard = gg_test_haz$est
  )
)


#********** Smoothed empirical hazard

std_dat  <- df_veteran %>% filter(trt == "Standard")
test_dat <- df_veteran %>% filter(trt == "Test")

mh_std <- muhaz(std_dat$time, std_dat$event)
mh_test <- muhaz(test_dat$time, test_dat$event)

haz_emp <- bind_rows(
  data.frame(
    Group = "Standard",
    time = mh_std$est.grid,
    hazard = mh_std$haz.est
  ),
  data.frame(
    Group = "Test",
    time = mh_test$est.grid,
    hazard = mh_test$haz.est
  )
)


#*********  Cumulative hazard plot

p1 <- ggplot() +
  geom_step(
    data = na_table,
    aes(x = time, y = cumhaz, color = Group, linetype = "Nelson-Aalen"),
    linewidth = 1
  ) +
  geom_line(
    data = gen_df,
    aes(x = time, y = cumhaz, color = Group, linetype = "GenGamma"),
    linewidth = 1
  ) +
  scale_color_manual(values = c("Standard" = "black", "Test" = "#D62728")) +
  scale_linetype_manual(values = c("Nelson-Aalen" = "solid", "GenGamma" = "dashed")) +
  labs(
    title = "Cumulative hazard",
    x = "Time (days)",
    y = "Cumulative hazard",
    color = "Treatment",
    linetype = "Curve"
  ) +
  theme_pubr(base_size = 12) +
  theme(
    plot.title = element_text(face = "bold", size = 13),
    legend.position = "bottom"
  )


#*********** Hazard plot

p2 <- ggplot() +
  geom_line(
    data = haz_emp,
    aes(x = time, y = hazard, color = Group, linetype = "Empirical"),
    linewidth = 1
  ) +
  geom_line(
    data = gen_df,
    aes(x = time, y = hazard, color = Group, linetype = "GenGamma"),
    linewidth = 1
  ) +
  scale_color_manual(values = c("Standard" = "black", "Test" = "#D62728")) +
  scale_linetype_manual(values = c("Empirical" = "solid", "GenGamma" = "dashed")) +
  labs(
    title = "Hazard",
    x = "Time (days)",
    y = "Hazard",
    color = "Treatment",
    linetype = "Curve"
  ) +
  theme_pubr(base_size = 12) +
  theme(
    plot.title = element_text(face = "bold", size = 13),
    legend.position = "bottom"
  )


combined_plot <- p1 + p2 +
  plot_layout(guides = "collect") &
  theme(
    legend.position = "right",
    legend.box = "vertical"
  )

combined_plot

ggsave("clean_hazard_comparison.png", combined_plot, width = 12, height = 5.5, dpi = 300)
```









# Step 3: Non-parametric hypothesis tests


## 3.1 Log rank test 


```{r logrank}
logrank_trt <- survdiff(surv_obj ~ trt, data = df_veteran, rho = 0)
logrank_trt

logrank_p <- 1 - pchisq(logrank_trt$chisq, df = length(logrank_trt$n) - 1)
logrank_p
```


```{r}

#********** Log-rank test table

logrank_trt <- survdiff(surv_obj ~ trt, data = df_veteran, rho = 0)
logrank_p <- 1 - pchisq(logrank_trt$chisq, df = length(logrank_trt$n) - 1)
logrank_df <- length(logrank_trt$n) - 1

logrank_tbl <- tibble(
  Group = sub("trt=", "", names(logrank_trt$n)),
  N = as.numeric(logrank_trt$n),
  Observed = as.numeric(logrank_trt$obs),
  Expected = round(as.numeric(logrank_trt$exp), 1),
  `(O-E)^2/E` = round((logrank_trt$obs - logrank_trt$exp)^2 / logrank_trt$exp, 4),
  `(O-E)^2/V` = round((logrank_trt$obs - logrank_trt$exp)^2 / logrank_trt$var[1,1], 5),
  `Chi-square` = round(logrank_trt$chisq, 3),
  `df` = logrank_df,
  `p-value` = round(logrank_p, 3)
)

logrank_kbl <- kbl(
  logrank_tbl,
  format = "html",
  align = "lcccccccc",
  caption = ""
) %>%
  kable_styling(
    full_width = FALSE,
    font_size = 18,
    bootstrap_options = c("striped", "hover", "condensed", "responsive")
  ) %>%
  row_spec(0, bold = TRUE, color = "white", background = "#1F4E79") %>%
  row_spec(1:nrow(logrank_tbl), extra_css = "border-bottom: 1px solid #999;") %>%
  column_spec(1, color = "black", width = "10em") %>%
  column_spec(2:9, color = "black", width = "8em")

logrank_kbl

save_kable(logrank_kbl, "logrank.png", zoom = 4)
```

## 3.2 Wilcoxon(Breslow) test

$$
P\!\left(\chi^2 \ge \chi_{\text{obs}}^2\right)
$$


```{r}
wilcoxon_trt <- survdiff(surv_obj ~ trt, data = df_veteran, rho = 1)
wilcoxon_trt

wilcoxon_p <- 1 - pchisq(wilcoxon_trt$chisq, df = length(wilcoxon_trt$n) - 1)
wilcoxon_p
```


```{r}
wilcoxon_trt <- survdiff(surv_obj ~ trt, data = df_veteran, rho = 1)
wilcoxon_p <- 1 - pchisq(wilcoxon_trt$chisq, df = length(wilcoxon_trt$n) - 1)
wilcoxon_df <- length(wilcoxon_trt$n) - 1

wilcoxon_tbl <- tibble(
  Group = sub("trt=", "", names(wilcoxon_trt$n)),
  N = as.numeric(wilcoxon_trt$n),
  Observed = round(as.numeric(wilcoxon_trt$obs), 1),
  Expected = round(as.numeric(wilcoxon_trt$exp), 1),
  `(O-E)^2/E` = round((wilcoxon_trt$obs - wilcoxon_trt$exp)^2 / wilcoxon_trt$exp, 3),
  `(O-E)^2/V` = round((wilcoxon_trt$obs - wilcoxon_trt$exp)^2 / wilcoxon_trt$var[1,1], 3),
  `Chi-square` = rep(round(wilcoxon_trt$chisq, 3), 2),
  `df` = rep(wilcoxon_df, 2),
  `p-value` = rep(round(wilcoxon_p, 3), 2)
)

wilcoxon_kbl <- kbl(
  wilcoxon_tbl,
  format = "html",
  align = "lcccccccc",
  caption = ""
) %>%
  kable_styling(
    full_width = FALSE,
    font_size = 18,
    bootstrap_options = c("striped", "hover", "condensed", "responsive")
  ) %>%
  row_spec(0, bold = TRUE, color = "white", background = "#1F4E79") %>%
  row_spec(1:nrow(wilcoxon_tbl), extra_css = "border-bottom: 1px solid #999;") %>%
  column_spec(1, color = "black", width = "10em") %>%
  column_spec(2:9, color = "black", width = "8em")

wilcoxon_kbl


save_kable(wilcoxon_kbl, "wilcoxon.png", zoom = 4)
```