---
title: "Group 1 Survival Analysis Project"
author: "Group 1"
date: "2026-03-22"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## STEP 1: Data Preparation and Baseline Table

```{r}
library(survival)
library(dplyr)
library(tidyr)
library(finalfit)
library(knitr)
library(kableExtra)
library(ggplot2)
library(survminer)
library(flexsurv)
library(gridExtra)
library(tibble)

data(cancer, package = "survival")

df_veteran <- veteran

if (all(sort(unique(na.omit(df_veteran$status))) == c(1, 2))) {
  df_veteran <- df_veteran %>%
    mutate(status = status - 1)
}

df_veteran <- df_veteran %>%
  mutate(
    trt = factor(trt, levels = c(1, 2), labels = c("Standard", "Test")),
    celltype = as.character(celltype),
    celltype = case_when(
      celltype %in% c("1", "squamous") ~ "Squamous",
      celltype %in% c("2", "smallcell") ~ "Small cell",
      celltype %in% c("3", "adeno") ~ "Adeno",
      celltype %in% c("4", "large") ~ "Large",
      TRUE ~ celltype
    ),
    celltype = factor(celltype, levels = c("Squamous", "Small cell", "Adeno", "Large")),
    prior = factor(ifelse(prior == 0, "No", "Yes"), levels = c("No", "Yes")),
    event_label = factor(status, levels = c(0, 1), labels = c("Censored", "Death"))
  ) %>%
  mutate(
    age = ff_label(age, "Age (years)"),
    karno = ff_label(karno, "Karnofsky score"),
    diagtime = ff_label(diagtime, "Months from diagnosis"),
    time = ff_label(time, "Survival time (days)"),
    celltype = ff_label(celltype, "Cell type"),
    prior = ff_label(prior, "Prior therapy"),
    event_label = ff_label(event_label, "Death (event)")
  )
```



## STEP 2: Exploratory Analysis




```{r}
dependent <- "trt"
explanatory <- c("age", "karno", "diagtime", "time", "celltype", "prior", "event_label")

table1 <- df_veteran %>%
  summary_factorlist(
    dependent = dependent,
    explanatory = explanatory,
    total_col = TRUE,
    p = TRUE,
    na_include = FALSE,
    p_cont_para = "t.test",
    p_cat = "fisher",
    add_dependent_label = FALSE
  ) %>%
  ff_row_totals(
    .data = df_veteran,
    dependent = dependent,
    explanatory = explanatory,
    total_name = "Total N",
    na_name = "Missing N"
  )

if ("label" %in% names(table1)) {
  table1 <- table1 %>% rename(Variable = label)
}

if ("all" %in% names(table1)) {
  table1 <- table1 %>% rename(Overall = all)
}

n_control <- sum(df_veteran$trt == "Standard", na.rm = TRUE)
n_test <- sum(df_veteran$trt == "Test", na.rm = TRUE)
n_total <- nrow(df_veteran)

names(table1)[names(table1) == "Standard"] <- paste0("Control (N=", n_control, ")")
names(table1)[names(table1) == "Test"] <- paste0("Test therapy (N=", n_test, ")")
names(table1)[names(table1) == "Overall"] <- paste0("Overall (N=", n_total, ")")

if ("Variable" %in% names(table1)) {
  table1 <- table1 %>%
    mutate(Variable = paste0(Variable, "   "))
}
baseline <- kbl(
  table1,
  format = "html",
  align = "l",
  caption = ""
) %>%
  kable_styling(
    full_width = FALSE,
    font_size = 18,
    bootstrap_options = c("striped", "hover", "condensed", "responsive")
  ) %>%
  row_spec(0, bold = TRUE, color = "white", background = "#1F4E79") %>%
  row_spec(1:nrow(table1), extra_css = "border-bottom: 1px solid #999;") %>%
  column_spec(1, color = "black", width = "8em") %>%
  column_spec(2, color = "black", width = "8em") %>%
  column_spec(3, color = "black", width = "8em") %>%
  column_spec(4, color = "black", width = "8em") %>%
  column_spec(5, color = "black", width = "8em") %>%
  column_spec(6, color = "black", width = "8em") %>%
  column_spec(7, color = "black", width = "8em") %>%
  column_spec(8, color = "black", width = "8em")

baseline

save_kable(baseline, "baseline_table.png", zoom = 4)
```






```{r}
library(knitr)
library(kableExtra)

sum_tab <- data.frame(
  "Sample Size" = nrow(veteran),
  "Events" = sum(veteran$status == 1),
  "Censored" = sum(veteran$status == 0),
  "Mean Time" = round(mean(veteran$time, na.rm = TRUE), 2),
  "Min Time" = min(veteran$time, na.rm = TRUE),
  "Max Time" = max(veteran$time, na.rm = TRUE),
  "Mean Age" = round(mean(veteran$age, na.rm = TRUE), 2),
  "Mean Karno" = round(mean(veteran$karno, na.rm = TRUE), 2),
  check.names = FALSE
)

sum_kbl <- kbl(
  sum_tab,
  format = "html",
  align = "c",
  caption = ""
) %>%
  kable_styling(
    full_width = FALSE,
    font_size = 18,
    bootstrap_options = c("striped", "hover", "condensed")
  ) %>%
  row_spec(0, bold = TRUE, color = "white", background = "#1F4E79") %>%
  row_spec(1:nrow(sum_tab), extra_css = "border-bottom: 1px solid #999;") %>%
  column_spec(1, color = "black", width = "8em") %>%
  column_spec(2, color = "black", width = "8em") %>%
  column_spec(3, color = "black", width = "8em") %>%
  column_spec(4, color = "black", width = "8em") %>%
  column_spec(5, color = "black", width = "8em") %>%
  column_spec(6, color = "black", width = "8em") %>%
  column_spec(7, color = "black", width = "8em") %>%
  column_spec(8, color = "black", width = "8em")

sum_kbl

save_kable(sum_kbl, "summary_table.png", zoom = 4)
```





### Minimum and Maximum 

```{r}
range_plot_data <- df_veteran %>%
  summarise(
    age_min = min(age, na.rm = TRUE),
    age_max = max(age, na.rm = TRUE),
    karno_min = min(karno, na.rm = TRUE),
    karno_max = max(karno, na.rm = TRUE),
    diagtime_min = min(diagtime, na.rm = TRUE),
    diagtime_max = max(diagtime, na.rm = TRUE),
    time_min = min(time, na.rm = TRUE),
    time_max = max(time, na.rm = TRUE)
  ) %>%
  pivot_longer(
    cols = everything(),
    names_to = c("Variable", "Statistic"),
    names_sep = "_",
    values_to = "Value"
  )

ggplot(range_plot_data, aes(x = Variable, y = Value, fill = Statistic)) +
  geom_col(position = position_dodge(width = 0.7), width = 0.6) +
  geom_text(
    aes(label = round(Value, 1)),
    position = position_dodge(width = 0.7),
    vjust = -0.3,
    size = 4
  ) +
  scale_fill_manual(values = c(
  "min" = "darkgreen",
  "max" = "#A52A2A"
))+
  labs(
    title = "Minimum and Maximum Values of Key Variables",
    x = "Variable",
    y = "Value"
  ) +
  theme_pubr(base_size = 13)+
  theme(legend.position = "right", plot.title = element_text(hjust = 0.1,vjust = 0.1, ))
```













### Survival times distribution




```{r}
#| warning: false
#| message: false


ggplot(df_veteran, aes(x = time)) +
  geom_histogram(binwidth = 40, fill = "darkgreen", color = "white") +
  labs(
    title = "Distribution of Survival Times",
    x = "Survival time (days)",
    y = "Number of patients"
  ) +
  scale_x_continuous(limits = c(0, max(df_veteran$time)), expand = c(0, 0)) +
  theme_pubr()+
  theme(legend.position = "right", plot.title = element_text(hjust = 0.2,vjust = 0.1, ))
```




### Swimmer plot to visualize the death and censored


```{r}
swimmer_df <- df_veteran %>%
  mutate(
    id = row_number(),
    start = 0,
    end = time,
    event = ifelse(status == 1, 1, 0)
  ) %>%
  slice(1:20)

swimmer_plot <- ggplot(swimmer_df, aes(y = factor(id))) +
  geom_segment(
    aes(x = start, xend = end, yend = factor(id)),
    linewidth = 1.2,
    colour = "steelblue"
  ) +
  geom_point(
    aes(x = end, shape = factor(event), colour = factor(event)),
    size = 3.5
  ) +
  scale_shape_manual(values = c(16, 17), labels = c("Censored", "Death")) +
  scale_colour_manual(values = c("grey40", "firebrick"), labels = c("Censored", "Death")) +
  labs(
    x = "Time (days)",
    y = "Patient ID",
    shape = "Outcome",
    colour = "Outcome",
    title = "Sample of censored and death"
  ) +
   theme_pubr()+
  theme(legend.position = "right", plot.title = element_text(hjust = 0.2,vjust = 0.1, ))

swimmer_plot
```


## STEP 3: Fit Parametric Models



```{r}

#************* creating an object to indicate censored with + sign(83+)

surv_obj <- Surv(time = df_veteran$time, event = df_veteran$status)
surv_obj
```

$$f(t)=\lambda e^{-\lambda t}, \quad t>0$$

$$\text{flexsurv uses MLE to estimate parameters}$$


```{r}
exp_model      <- flexsurvreg(surv_obj ~ 1, data = df_veteran, dist = "exp")
wei_model      <- flexsurvreg(surv_obj ~ 1, data = df_veteran, dist = "weibull")
ln_model       <- flexsurvreg(surv_obj ~ 1, data = df_veteran, dist = "lnorm")
ll_model       <- flexsurvreg(surv_obj ~ 1, data = df_veteran, dist = "llogis")
gom_model      <- flexsurvreg(surv_obj ~ 1, data = df_veteran, dist = "gompertz")
gamma_model    <- flexsurvreg(surv_obj ~ 1, data = df_veteran, dist = "gamma")
gengamma_model <- flexsurvreg(surv_obj ~ 1, data = df_veteran, dist = "gengamma")

```




### AIC, LogLik, and MLE Estimates




```{r}
library(dplyr)
library(knitr)
library(kableExtra)

get_mle_parameters_estimate <- function(model) {
  est <- model$res[, "est", drop = FALSE]
  param_names <- rownames(est)

  # Rename parameter names to symbols
  param_names[param_names == "mu"] <- "&mu;"
  param_names[param_names == "sigma"] <- "&sigma;"
  param_names[param_names == "shape"] <- "k"
  param_names[param_names == "scale"] <- "&lambda;"
  param_names[param_names == "rate"] <- "&lambda;"
  param_names[param_names == "meanlog"] <- "&mu;<sub>log</sub>"
  param_names[param_names == "sdlog"] <- "&sigma;<sub>log</sub>"

  paste(param_names, "=", round(est[, 1], 4), collapse = "<br>")
}

#********** Function to display Generalized Gamma in lecture-note form a, d, p

get_gengamma_adp <- function(model) {
  mu <- model$res["mu", "est"]
  sigma <- model$res["sigma", "est"]
  Q <- model$res["Q", "est"]

  k <- 1 / (Q^2)
  p <- Q / sigma
  a <- exp(mu - log(k) / p)
  d <- p * k

  paste0(
    "a = ", round(a, 4), "<br>",
    "d = ", round(d, 4), "<br>",
    "p = ", round(p, 4)
  )
}

model_table <- data.frame(
  Model = c(
    "Exponential",
    "Weibull",
    "Log-normal",
    "Log-logistic",
    "Gompertz",
    "Gamma",
    "Generalized Gamma"
  ),
  LogLik = c(
    exp_model$loglik,
    wei_model$loglik,
    ln_model$loglik,
    ll_model$loglik,
    gom_model$loglik,
    gamma_model$loglik,
    gengamma_model$loglik
  ),
  AIC = c(
    AIC(exp_model),
    AIC(wei_model),
    AIC(ln_model),
    AIC(ll_model),
    AIC(gom_model),
    AIC(gamma_model),
    AIC(gengamma_model)
  ),
  "MLE Estimates" = c(
    get_mle_parameters_estimate(exp_model),
    get_mle_parameters_estimate(wei_model),
    get_mle_parameters_estimate(ln_model),
    get_mle_parameters_estimate(ll_model),
    get_mle_parameters_estimate(gom_model),
    get_mle_parameters_estimate(gamma_model),
    get_gengamma_adp(gengamma_model)
  ),
  stringsAsFactors = FALSE,
  check.names = FALSE
)

model_table <- model_table %>%
  mutate(
    LogLik = round(LogLik, 4),
    AIC = round(AIC, 4)
  )

modeltable <- kbl(
  model_table,
  format = "html",
  escape = FALSE,
  align = c("l", "c", "c", "l"),
  caption = ""
) %>%
  kable_styling(
    full_width = FALSE,
    font_size = 14,
    bootstrap_options = c("striped", "hover", "condensed")
  ) %>%
  row_spec(0, bold = TRUE, color = "white", background = "#1F4E79") %>%
  row_spec(1:nrow(model_table), extra_css = "border-bottom: 1px solid #999;") %>%
  column_spec(1, color = "black", width = "12em") %>%
  column_spec(2, color = "black", width = "8em") %>%
  column_spec(3, color = "black", width = "8em") %>%
  column_spec(4, color = "black", width = "22em")

modeltable

save_kable(modeltable, "modeltable.png", zoom = 4)

```





## STEP 4: Model Selection and Survival Curve Comparison


```{r}
best_model_name <- model_table$Model[which.min(model_table$AIC)]

best_fit <- switch(
  best_model_name,
  "Exponential" = exp_model,
  "Weibull" = wei_model,
  "Log-normal" = ln_model,
  "Log-logistic" = ll_model,
  "Gompertz" = gom_model,
  "Gamma" = gamma_model,
  "Generalized Gamma" = gengamma_model
)

best_model_summary <- tibble(
  `Best Model` = best_model_name,
  LogLik = round(model_table$LogLik[which.min(model_table$AIC)], 4),
  AIC = round(min(model_table$AIC), 4)
)

best_model_table<- kbl(
  best_model_summary,
  format = "html",
  align = "c",
  caption = ""
) %>%
  kable_styling(full_width = FALSE, bootstrap_options = c("striped", "hover", "condensed")) %>%
  row_spec(0, bold = TRUE, color = "white", background = "#1F4E79") %>%
  add_header_above(
    background = "#2E86C1",
    color = "white",
    bold = TRUE
  )%>%
  row_spec(1:nrow(best_model_summary), extra_css = "border-bottom: 1px solid #999;") %>%
  column_spec(1, color = "black", width = "12em") %>%
  column_spec(2, color = "black", width = "8em") %>%
  column_spec(3, color = "black", width = "8em")


save_kable(best_model_table, "best_model_table.png", zoom = 4)
```



 
## Survival curves for each models  and kaplan curve  





```{r}





t_grid <- seq(0, max(df_veteran$time), length.out = 200)

#************* Kaplan-Meier fit

km_fit <- survfit(Surv(time, status) ~ 1, data = df_veteran)

km_data <- data.frame(
  time = km_fit$time,
  surv = km_fit$surv,
  Model = "Kaplan-Meier"
)


#************ Function to extract survival curve from each model
get_surv_df <- function(model, model_name) {
  s <- summary(model, type = "survival", t = t_grid)
  data.frame(
    time = s[[1]]$time,
    surv = s[[1]]$est,
    Model = model_name,
    Comparison = model_name
  )
}


model_curves <- bind_rows(
  get_surv_df(exp_model, "Exponential"),
  get_surv_df(wei_model, "Weibull"),
  get_surv_df(ln_model, "Log-normal"),
  get_surv_df(ll_model, "Log-logistic"),
  get_surv_df(gom_model, "Gompertz"),
  get_surv_df(gamma_model, "Gamma"),
  get_surv_df(gengamma_model, "Generalized Gamma")
)

km_panels <- bind_rows(
  lapply(unique(model_curves$Comparison), function(m) {
    km_data %>% mutate(Comparison = m)
  })
)

plot_data <- bind_rows(km_panels, model_curves)

surv_plot <- ggplot(plot_data, aes(x = time, y = surv, color = Model, linetype = Model)) +
    geom_step(
    data = subset(plot_data, Model == "Kaplan-Meier"),
    linewidth = 1.2
  ) +
  geom_line(
    data = subset(plot_data, Model != "Kaplan-Meier"),
    linewidth = 1.2
  ) + 
  facet_wrap(~ Comparison, ncol = 2, axes = "all_x", axis.labels = "all_x") +
  scale_color_manual(values = c(
    "Kaplan-Meier" = "black",
    "Exponential" = "#E41A1C",
    "Weibull" = "#377EB8",
    "Log-normal" = "#4DAF4A",
    "Log-logistic" = "#984EA3",
    "Gompertz" = "#FF7F00",
    "Gamma" = "#A65628",
    "Generalized Gamma" = "#F781BF"
  )) +
  labs(
    title = "Kaplan-Meier Compared with Each Parametric Model",
    x = "Time (days)",
    y = "Survival probability",
    color = "Curve",
    linetype = "Curve"
  ) +
  theme_classic() +
  theme(
    strip.background = element_blank(),
    panel.border = element_blank(),
    strip.text = element_text(face = "bold", size = 8),
    legend.position = "right",
    plot.title = element_text(hjust = 0.4, vjust = 0.1),
    legend.text = element_text(size = 8),
    legend.title = element_text(size =8),
    axis.text = element_text(size = 8),
    axis.title = element_text(size = 8)
  )

surv_plot


ggsave(
  filename = "survival_model_comparison.png")
```





## STEP 5: Life Functions for the Best Model



```{r}
t <- seq(0, max(df_veteran$time), length.out = 200)

S <- summary(best_fit, t = t, type = "survival")[[1]]$est
h <- summary(best_fit, t = t, type = "hazard")[[1]]$est
H <- summary(best_fit, t = t, type = "cumhaz")[[1]]$est
f <- h * S
F <- 1 - S
```



### Mean and variance 


```{r}
library(tibble)
library(knitr)
library(kableExtra)

mu <- gengamma_model$res["mu", "est"]
sigma <- gengamma_model$res["sigma", "est"]
Q <- gengamma_model$res["Q", "est"]

k <- 1 / Q^2
p <- Q / sigma
a <- exp(mu - log(k) / p)
d <- p * k

mean_T <- a * gamma((d + 1)/p) / gamma(d/p)

var_T <- a^2 * (
  gamma((d + 2)/p) / gamma(d/p) -
    (gamma((d + 1)/p) / gamma(d/p))^2
)

gengamma_summary <- tibble(
  Model = "Generalized Gamma",
  a = round(a, 4),
  d = round(d, 4),
  p = round(p, 4),
  Mean= round(mean_T, 4),
  Variance = round(var_T, 4)
)

gengamma_summary<-gengamma_summary %>%
  kbl(
    format = "html",
    align = "c",
    caption = ""
  ) %>%
  kable_styling(
    full_width = FALSE,
    bootstrap_options = c("striped", "hover", "condensed")
  ) %>%
  row_spec(0, bold = TRUE, color = "white", background = "#1F4E79") %>%
  row_spec(1:nrow(gengamma_summary), extra_css = "border-bottom: 1px solid #999;") %>%
  column_spec(1, color = "black", width = "12em") %>%
  column_spec(2, color = "black", width = "8em") %>%
  column_spec(3, color = "black", width = "8em") %>%
  column_spec(4, color = "black", width = "8em") %>%
  column_spec(5, color = "black", width = "12em") %>%
  column_spec(6, color = "black", width = "12em")


save_kable(gengamma_summary, "genvariance.png", zoom=4)
```












